Title on one or two lines without capitals, except after a colon

### 2020-06-16 Report/Sections/background.tex
Revise Report

$\delta$ is often assumed to be the stationary distribution of the $\Gamma$- an assumption that we make in this work. Further, the distribution of each observation $Y_t$ depends only on the value of $X_t$ and none of the preceding observations or behavioral states, i.e. $p(y_t|x_1,\ldots, x_N, y_1,\ldots,y_{t-1},y_{t+1},\ldots,y_N) = p(y_t|x_t)$. In particular, we denote the distribution of $Y_t$ given $X_t = s$ as $f_s(y_t)$. $f_1,\ldots,f_N$ in turn depend upon a vector of parameters $\Theta = (\theta_1,\ldots,\theta_N)$ and hereafter are referred to as ``emission distributions". A visualization of the dependence structure can be seen in figure (\ref{fig:HMM}).
In order to estimate the parameters of an HMM, suppose data $y = (y_1,\ldots,y_T)$ is observed as a realization of an unknown HMM model. The probability transition matrix $\Gamma$, and parameters of the emission distributions $\Theta$ can be estimated by maximizing the likelihood of $y$, which we denote as $\calL_{\text{HMM}}(y;\Theta,\Gamma)$, with respect to $\Gamma$, $\delta$, and $\Theta$. In addition, $\calL_{\text{HMM}}(y;\theta,\Gamma,\delta)$ can be calculated using the \textit{forward algorithm} \cite{Zucchini:2016} as follows:
%
$$\calL_{\text{HMM}}(y;\theta,\Gamma) = \delta P(y_1;\theta) \prod_{t=2}^T \Gamma P(y_t;\theta) \mathbf{1}$$
%
where $\mathbf{1}_N$ is an $N$-dimensional column vector of ones, and:
%
$$P(y_t;\theta) = \text{diag}(f_1(y_t),\ldots , f_N(y_t)).$$
%
In order to ensure remove constraints when maximizing $\calL_{\text{HMM}}(y)$, $\Gamma$ is often parameterized using $\eta \in \bbR^{N \times N}$ and the following link function:
%
$$\Gamma_{ij} = \frac{\exp(\eta_{ij})}{\sum_{k=1}^N \exp(\eta_{ik})}, \qquad \eta_{ii} = 0 \quad \forall i \in \{1, \ldots, N\}.$$
%
$\eta_{ii}$ is set to zero for identifiability. $\calL_{\text{HMM}}(y;\theta,\Gamma)$ can be maximized using any numerical optimizer.

@@

A hierarchical hidden Markov model (or HHMM) is a variation of a hidden Markov model in which $X_t$ satisfies the conditions of section \ref{subsection:HMMs}, but in addition it produces yet another fine-scale hidden Markov model of length $T_t^*$. This fine-scale hidden Markov model can be either instead of \cite{Barajas:2017} or in addition to \cite{Adam:2019} a coarse-scale observation $Y_t$ which must satisfy the conditions of section \ref{subsection:HMMs}. The fine-scale HMM is comprised of a fine-scale Markov chain $X^*_t = (X^*_{t,1}, \ldots, X^*_{t,T^*_t})$ and fine-scale observations $Y^*_t = (Y^*_{t,1}, \ldots, Y^*_{t,T^*_t})$, each of length $T^*_t$. $X^*_t$ can take one of $N^*$ values. Note that while $N^*$ can be set to depend upon $X_t$, we assume here that it does not for simplicity in both the model and the notation. $X^*_t$ is characterized by an initial distribution $\delta^{*(X_t)} \in \bbR^{N^*}$ and probability transition matrix $\Gamma^{*(X_t)} \in \bbR^{N^* \times N^*}$:
%
$$\delta^{*(s)}_i = Pr(X^*_{t,1} = i | X_t = s); \quad i = 1,\ldots,N^*$$
%
$$\Gamma^{*(s)}_{ij} = Pr(X^*_{t,t^*+1} = j | X^*_{t,t^*} = i, X_t = s); \quad t^* = 1, \ldots, T^*_t-1; \enspace i,j = 1,\ldots,N^*$$
%
Again, we assume that the initial distribution $\delta^{*(s)}$ is the stationary distribution of $\Gamma^{*(s)}$. Each fine-scale observation $Y^*_{t,t^*}$ depends only on the value of its corresponding fine-scale hidden state $X^*_{t,t^*}$ and coarse-scale hidden state, $X_t$. We denote the distribution of $Y^*_{t,t^*}$ given that $X_t = s$ and $X^*_{t,t^*} = s^*$ as $f_{s,s^*}$. Therefore, in the hierarchical setting $\Theta$ is a $N$-by-$N^*$ matrix of parameters where $\theta_{s,s^*}$ describes the distribution $f_{s,s^*}$. Note the parameters of the fine-scale hidden Markov model ($\Gamma^{*(X_t)}$, $\delta^{*(X_t)}$, and $\theta^{*(X_t)}$) all depend upon the hidden state of the \textit{crude-scale} hidden Markov model. However, depending upon the discretion of the researcher, it is possible to force any of these parameters to be independent of the crude-scale hidden state $X_t$. A visualization of the full structure of the HHMM can be seen in figure (\ref{fig:HHMM}). Due to the nested structure of a hierarchical hidden Markov model, the likelihood of an HHMM is still easy to calculate using the forward algorithm:
%
$$\calL_{\text{HHMM}}(y,y^*;\theta,\theta^*,\Gamma,\Gamma^*,\delta,\delta^*) = \delta P(y_1,y_1^*;\theta,\theta^*,\Gamma^*,\delta^*) \prod_{t=2}^T \Gamma P(y_t,y_t^*;\theta,\theta^*,\Gamma^*,\delta^*) \mathbf{1}$$
%
where:
%
\begin{align*}
	P(y_t,y_t^*;\theta,\theta^*,\Gamma^*,\delta^*)  = \text{diag}\left[p_{\theta}(y_t|x_t = x_1)\calL_{\text{HMM}}\left(y_t^*;\theta^{*(x_1)},\Gamma^{*(x_1)},\delta^{*(x_1)}\right), . . . , \right.\\
	\left. p_{\theta}(y_t|x_t = x_N )\calL_{\text{HMM}}\left(y_t^*;\theta^{*(x_N)},\Gamma^{*(x_N)},\delta^{*(x_N)}\right) \right]
\end{align*}
%
Note that this formulation assumes that the crude-scale observations at a given time $Y_t$ and the fine-scale observation time series $Y_t^*$ are independent of one another when conditioned on $X_t$. For more information on specific considerations for HHMMs such as incorporating covariates into the probability transition matrix, model selection and model checking, see Adam et al \cite{Adam:2019}.

@@

\subsection{Conditionally auto-regressive HMMs}

One of the key assumptions of both HMMs and HHMMs is \textit{conditional Independence} between observations at both the crude and fine scale. Namely, given the state $X_t$ or $X^*_{t,t^*}$, $Y_t$ or $Y^*_{t,t^*}$ (respectively) is assumed to be independent from all other observations. Therefore, traditional HMMs and HHMMs can fail when the observation sequence $Y$ exhibits significant auto-correlation in time. Examples include fluking in marine mammals in Vancouver, BC (see the results section) and the swimming behavior of horn sharks off the coast of Southern California \cite{Adam:2019}.

One way to deal with auto-correlation in fine-scale behavioral processes is to use the CarHMM, or \textit{conditionally auto-regressive hidden Markov model}, introduced by Lawler et al \cite{Lawler:2019}. If the emission distribution of observation $Y_t$ is parameterized by its mean and variance, i.e. $\theta_{x_t} = \{\mu_{x_t},\sigma^2_{x_t}\}$, then the CarHMM models the distribution of $Y_t|X_t,Y_{t-1}$ as the following:
%
$$\mathbb{E}(Y_t) = \phi_{x_t}*Y_{t-1} + (1-\phi_{x_t}) * \mu_{x_t}$$
$$\mathbb{V}(Y_t) = \sigma^2_{x_t}$$
%
The CarHMM therefore explicitly models auto-correlation into the emission distributions of the HMM while maintaining the structure needed to run the forward algorithm and only adding one additional parameter per observation and hidden state.

The likelihood of CarHMM is still compatible with the forward algorithm:
\begin{equation}
\calL_{\text{CarHMM}}(y) = \delta \prod_{t=2}^T \Gamma P(y_t;\theta) \mathbf{1}
\label{CarHMM_likelihood}
\end{equation}
where:
%
$$P(y_t;\theta) = \text{diag}(p_\theta(y_t|y_{t-1},X_t = x_1), . . . , p_\theta(y_t|y_{t-1},X_t = x_N )), \qquad t > 1$$
%
and the first observation $y_1$ is assumed to be fixed as an initial value. The graphical model associated with the structure of a CarHMM is shown in figure (\ref{fig:CarHMM}).

@@

Once an HMM, HHMM, or CarHMM model is fit by maximizing their respective likelihoods with respect to their parameters, it is common to find the most likely sequence of hidden states $\hat X$.

This can be done in one of two ways: first, the \textit{most likely sequence} of \text{all} hidden states $\hat X$ is often found using a dynamic programming algorithm called the Viterbi algorithm \cite{Viterbi:1967}. Second, the \textit{probability} of \textit{each} crude-level state (conditioned on the learned parameters) can be found using the \textit{forward-backward algorithm}. In the case of HHMMs, both the Viterbi algorithm and the forward-backward algorithm are easy to extend to HHMMs by recursively running the algorithm on the fine-scale HMMs after decoding the states of the coarse-scale HMM.

While both the Viterbi algorithm and the forward-backward algorithm are used in the current ecology literature, we use the \textit{forward-backward algorithm} in this work for several reasons. In particular, the forward-backward algorithm is used to find the \textit{psuedoresiduals} of a given model, which is an important tool for model validation. In addition, for HHMMs the forward-backward algorithm can be used recursively to find the probability of the fine-level states $X^*_{t^*,t}$ exactly by marginalizing out $X_t$:
%
$$P(X^*_{t^*,t} = x^*_{t^*,t}) = \sum_{n=1}^N P(X_t = x_n)P(X^*_{t^*,t} = x^*_{t^*,t} | X_t = x_n)$$
%
where $P(X_t = x_n)$ can be found using the forward-backward algorithm on the crude-level Markov chain and $P(X^*_{s,t} = x^*_{s,t} | X_t = x_n)$ can be found by running the forward-backward algorithm on the fine-level HMM for every possible value of $X_t$.

>>>>

$\delta$ is often assumed to be the stationary distribution of the $\Gamma$- an assumption that we make in this work. Further, the distribution of each observation $Y_t$ depends only on the value of $X_t$ and none of the preceding observations or behavioral states, i.e. $p(y_t|x_1,\ldots, x_N, y_1,\ldots,y_{t-1},y_{t+1},\ldots,y_N) = p(y_t|x_t)$. In particular, we denote the distribution of $Y_t$ given $X_t = s$ as $f^{(s)}(y_t)$. $f^{(1)},\ldots,f^{(N)}$ in turn depend upon a vector of parameters $\Theta = (\theta^{(1)},\ldots,\theta^{(N)})$ and hereafter are referred to as ``emission distributions". A visualization of the dependence structure can be seen in figure (\ref{fig:HMM}).
In order to estimate the parameters of an HMM, suppose data $y = (y_1,\ldots,y_T)$ is observed as a realization of an unknown HMM model. The probability transition matrix $\Gamma$, and parameters of the emission distributions $\Theta$ can be estimated by maximizing the likelihood of $y$, which we denote as $\calL_{\text{HMM}}(y;\Theta,\Gamma)$, with respect to $\Gamma$, $\delta$, and $\Theta$. In addition, $\calL_{\text{HMM}}(y;\theta,\Gamma,\delta)$ can be calculated using the \textit{forward algorithm} \cite{Zucchini:2016} as follows:
%
$$\calL_{\text{HMM}}(y;\theta,\Gamma) = \delta P(y_1;\theta) \prod_{t=2}^T \Gamma P(y_t;\theta) \mathbf{1}$$
%
where $\mathbf{1}_N$ is an $N$-dimensional column vector of ones, and:
%
$$P(y_t;\theta) = \text{diag}(f^{(1)}(y_t),\ldots , f^{(N)}(y_t)).$$
%
In order to ensure remove constraints when maximizing $\calL_{\text{HMM}}(y)$, $\Gamma$ is often parameterized using $\eta \in \bbR^{N \times N}$ and the following link function:
%
$$\Gamma_{ij} = \frac{\exp(\eta_{ij})}{\sum_{k=1}^N \exp(\eta_{ik})}, \qquad \eta_{ii} = 0 \quad \forall i \in \{1, \ldots, N\}.$$
%
$\eta_{ii}$ is set to zero for identifiability. $\calL_{\text{HMM}}(y;\theta,\Gamma)$ can be maximized using any numerical optimizer.

@@

A hierarchical hidden Markov model (or HHMM) is a variation of a hidden Markov model in which $X_t$ satisfies the conditions of section \ref{subsection:HMMs}, but in addition it produces yet another fine-scale hidden Markov model of length $T_t^*$. This fine-scale hidden Markov model can be either instead of \cite{Barajas:2017} or in addition to \cite{Adam:2019} a coarse-scale observation $Y_t$ which must satisfy the conditions of section \ref{subsection:HMMs}. The fine-scale HMM is comprised of a fine-scale Markov chain $X^*_t = (X^*_{t,1}, \ldots, X^*_{t,T^*_t})$ and fine-scale observations $Y^*_t = (Y^*_{t,1}, \ldots, Y^*_{t,T^*_t})$, each of length $T^*_t$. $X^*_t$ can take one of $N^*$ values. Note that while $N^*$ can be set to depend upon $X_t$, we assume here that it does not for simplicity in both the model and the notation. Conditioned on the coarse-scale hidden state $X_t = s$, $X^*_t$ is characterized by an initial distribution $\delta^{*(s)} \in \bbR^{N^*}$ and probability transition matrix $\Gamma^{*(s)} \in \bbR^{N^* \times N^*}$:
%
$$\delta^{*(s)}_{i^*} = Pr(X^*_{t,1} = i^* | X_t = s); \quad i^* = 1,\ldots,N^*$$
%
$$\Gamma^{*(s)}_{i^*j^*} = Pr(X^*_{t,t^*+1} = j^* | X^*_{t,t^*} = i^*, X_t = s); \quad t^* = 1, \ldots, T^*_t-1; \enspace i^*,j^* = 1,\ldots,N^*$$
%
Again, although not necessary, we assume that the initial distribution $\delta^{*(s)}$ is the stationary distribution of $\Gamma^{*(s)}$. In total, there are $N$ different values that $\Gamma^{*(s)}$ and $\delta^{*(s)}$ can take- one for each possible value of $X_t$. We denote the sets of all fine-scale initial distributions and probability transition matrices as $\delta^*$ and $\Gamma^*$, respectively, where $|\delta^*| = |\Gamma^*| = N$.

Each fine-scale observation $Y^*_{t,t^*}$ depends only on the value of its corresponding fine-scale hidden state $X^*_{t,t^*}$ and coarse-scale hidden state, $X_t$. In particular, $p(y^*_{t,t^*} | y^*_{t,1}, \ldots y^*_{t,t^*-1},y^*_{t,t^*+1},\ldots,y^*_{t,T^*_t},x^*_t,x,y) = p(y^*_{t,t^*}|x^*_{t,t^*},x_t)$. We denote the distribution of $Y^*_{t,t^*}$ given that $X_t = s$ and $X^*_{t,t^*} = s^*$ as $f^{*(s,s^*)}(y^*_{t,t^*})$. Conditioned on the fact that $X_t = s$, the fine-scale distributions $f^{*(s,1)}, \ldots, f^{*(s,N^*)}$ are parameterized by an $N^*$-dimensional vector of parameters $\Theta^{*(s)} = (\theta^{*(s,1)},\ldots,\theta^{*(s,N^*)})$, where $\theta^{*(s,s^*)}$ describes the distribution $f^{*(s,s^*)}$. Depending upon the behaviour and responses being modelled, it is possible to force the fine-scale emission distributions to be equal across all values of $s \in \{1, \ldots, N\}: \enspace \theta^{*(1,s^*)} = \theta^{*(2,s^*)} = \ldots = \theta^{*(N,s^*)}$. A visualization of the full structure of the HHMM can be seen in figure (\ref{fig:HHMM}).

Due to the nested structure of a hierarchical hidden Markov model, the likelihood of an HHMM is still easy to calculate using the forward algorithm. Suppose that both coarse-scale and fine scale observations ($y$ and $y^*$, respectively) are observed from an HHMM with unknown parameters. Then the likelihood of the observed data is:
%
$$\calL_{\text{HHMM}}(y,y^*;\Theta,\Theta^*,\Gamma,\Gamma^*) = \delta P(y_1,y_1^*;\Theta,\Theta^*,\Gamma^*) \prod_{t=2}^T \Gamma P(y_t,y_t^*;\Theta,\Theta^*,\Gamma^*) \mathbf{1}_N$$
%
where:
%
\begin{align*}
	P(y_t,y_t^*;\Theta,\Theta^*,\Gamma^*) = \text{diag}\left[&f^{(1)}(y_t)\calL_{\text{HMM}}\left(y_t^*;\Theta^{*(1)},\Gamma^{*(1)}\right),\ldots, \right. \\
	\left. &f^{(N)}(y_t)\calL_{\text{HMM}}\left(y_t^*;\Theta^{*(N)},\Gamma^{*(N)}\right) \right]
\end{align*}
%
Note that this formulation assumes that the coarse-scale observations at a given time $Y_t$ and the fine-scale observation time series $Y_t^*$ are independent of one another when conditioned on $X_t$. For more information on specific considerations for HHMMs such as incorporating covariates into the probability transition matrix, model selection and model checking, see Adam et al \cite{Adam:2019}.

@@

\subsection{conditionally auto-regressive HMMs}

One of the key assumptions of both HMMs and HHMMs is \textit{conditional independence} between observations at both the coarse and fine scale. Namely, given the state $X_t$, $Y_t$ is assumed to be independent from all other observations. Therefore, traditional HMMs and HHMMs can fail when the observations $Y$ exhibit significant correlation in time. Examples include fluking in marine mammals in Vancouver, BC (see the results section) and the swimming behavior of horn sharks off the coast of Southern California \cite{Adam:2019}.

One way to deal with auto-correlation in fine-scale behavioral processes is to use the CarHMM, or \textit{conditionally auto-regressive hidden Markov model}, introduced by Lawler et al \cite{Lawler:2019}. In particular, the CarHMM describes the distribution of $(Y_t|Y_{t-1}, X_t=s)$ with parameters $\theta^{(s)} = \{\mu^{(s)},\sigma^{2(s)},\phi^{(s)}\}$, where:
%
$$\mathbb{E}(Y_t|Y_{t-1} = y_{t-1},X_t=s) = \phi^{(s)}*y_{t-1} + (1-\phi^{(s)}) * \mu^{(s)}$$
$$\mathbb{V}(Y_t| X_t = s) = \sigma^{2(s)}$$
%
The CarHMM therefore explicitly models auto-correlation into the emission distributions of the HMM while maintaining the structure needed to run the forward algorithm and only adding one additional parameter per hidden state $(\phi^{(s)})$.

The likelihood of CarHMM is still compatible with the forward algorithm:
\begin{equation}
\calL_{\text{CarHMM}}(y) = \delta \prod_{t=2}^T \Gamma P(y_t;\Theta) \mathbf{1}_N
\label{CarHMM_likelihood}
\end{equation}
where:
%
$$P(y_t;\Theta) = \text{diag}\left(f^{(1)}(y_t|y_{t-1}), \ldots , f^{(N)}(y_t|y_{t-1}) \right)$$
%
and the first observation $y_1$ is assumed to be fixed as an initial value. The graphical model associated with the structure of a CarHMM is shown in figure (\ref{fig:CarHMM}).

@@

Once an HMM, HHMM, or CarHMM model is fit by maximizing its respective likelihood with respect to its parameters, it is common to find the most likely sequence of hidden states $\hat X$.

This can be done in one of two ways: first, the most likely \textit{sequence} of \text{all} hidden states can be found using a dynamic programming algorithm called the Viterbi algorithm \cite{Viterbi:1967}. Alternatively, the probability of \textit{each} coarse-level state (conditioned on the learned parameters) can be found using the \textit{forward-backward algorithm}. In the case of HHMMs, both the Viterbi algorithm and the forward-backward algorithm are easy to extend to HHMMs by recursively running the algorithm on the fine-scale HMMs after decoding the states of the coarse-scale HMM.

While both the Viterbi algorithm and the forward-backward algorithm are used in the current ecology literature, we use the \textit{forward-backward algorithm} in this work for several reasons. In particular, the forward-backward algorithm is used to find the \textit{psuedoresiduals} of a given model, which are an important tool for model validation. In addition, the forward-backward algorithm can be implemented recursively to find the probability of the fine-level states $X^*_{t^*,t}$ exactly by marginalizing out $X_t$:
%
$$P(X^*_{t^*,t} = s^*) = \sum_{s=1}^N P(X_t = s)P(X^*_{t^*,t} = s^* | X_t = s)$$
%
where $P(X_t = s)$ can be found using the forward-backward algorithm on the coarse-level Markov chain and $P(X^*_{s,t} = s^* | X_t = s)$ can be found by running the forward-backward algorithm on the fine-level HMM for every possible value of $s = 1, \ldots, N$.

### 2020-06-16 Report/Sections/carhmm_ou.tex
Revise Report

Although the CarHMM can incorporate auto-correlation into the structure of an HMM, they break down when observations are taken at irregular time intervals since the probability transition matrix $\Gamma$ would then depend upon the time step between observations. A common solution to this issue is to use a continuous-time method such as the one introduced by Michelot et al \cite{Michelot:2019}, which models the movement of an animal as an Ornstein-Uhlenbeck process with parameters that depend upon the underlying behavioural state of the animal. Namely, a state-switching Ornstein-Uhlenbeck process $y$ is the solution to the following stochastic differential equation:
%
$$dy_t = \beta_{x_t}(\gamma_{x_t} - y_t)dt + \omega_{x_t} dW_t$$
%
where $x_t$ is the fine-scale behavior of the animal at time $t$, $\beta_{x_t}$ relates to rate at which the process returns to its mean value, $\gamma_{x_t}$ is the long-term mean value of the process, $\omega_{x_t}$ is related to short-term variance, and $W$ is a Wiener process. $x_t$ is assumed to follow an unobserved continuous-time Markov process. Here, $t \in \mathbb{R}$ is a continuous time stamp rather than an index. If the behavioral state $x_t$ is known and does not change between observations, the solution is known to be the following \cite{Michelot:2019}:
%
\begin{align}
y_{t+\Delta t} | x_{t:t+\Delta t} \sim \mathcal{N}\left((1-e^{-\beta_{x_t}\Delta t})\gamma_{x_t} + e^{-\beta_{x_t}\Delta t} y_t,\quad \frac{\omega_{x_t}^2}{2\beta_{x_t}} (1-e^{-2\beta_{x_t}\Delta t})\right)
\label{eqn:OU_sol}
\end{align}
%
where $\Delta t$ is the distance between any two observations $Y_t$ and $Y_{t+\delta t}$. Continuous time models can be built up from arbitrarily complex stochastic differential equations and allow for uneven step lengths in the observations sequence $Y$. However, most continuous time models require MCMC algorithms to perform inference over and as a result are not easily incorporated into the HHMM structure.

@@

Once the CarHMM is fit by maximizing the likelihood in equation (\ref{CarHMM_likelihood}), the emission distributions can be interpreted as the solution of a state-switching Ornstein-Uhlenbeck process similar to the one introduced by Michelot et al \cite{Michelot:2019}. In order for this to be the case, however, the following two conditions must be met:
\begin{enumerate}
	\item The underlying behavioral state of the continuous-time model follows a Markov chain rather than a Markov process.
	\item The emission distributions of the CarHMM are normal.
\end{enumerate}
If this is the case, then the CarHMM is equivalent to a state-switching Ornstein-Uhlenbeck process. This gives new interpretation to the learned parameters of the CarHMM in the context of a continuous-time model.

\subsubsection{Proof}

Suppose that $\Delta t$ is constant for all observations. In addition, introduce the following transformations:
%
\begin{align}
\mu_{x_t} = \gamma_{x_t}, \qquad \phi_{x_t} = e^{-\beta_{x_t}\Delta t}, \qquad \sigma^2_{x_t} = \frac{\omega_{x_t}^2}{2\beta_{x_t}} (1-e^{-2\beta_{x_t}\Delta t}) \label{eqn:CarHMM_to_OU} \\
\gamma_{x_t} = \mu_{x_t}, \qquad \beta_{x_t} = -\frac{\log(\phi_{x_t})}{\Delta t}, \qquad \omega_{x_t} = \sqrt{\frac{2\sigma^{2}_{x_t}\log(\phi_{x_t})}{\Delta t (\phi_{x_t}^2-1)}}
\label{eqn:OU_to_CarHMM}
\end{align}
%
Combining equation (\ref{eqn:CarHMM_to_OU}) with equation (\ref{eqn:OU_sol}) gives the following:
%
\begin{align*}
y_{t+\Delta t} | x_{t:t+\Delta t} \sim \mathcal{N}\left((1-\phi_{x_t})\mu_{x_t} + \phi_{x_t} y_t,\quad \sigma_{x_t}^2 \right)
\end{align*}
%
If $\Delta t$ is fixed and $x_t$ is adjusted to follow a Markov chain rather than a Markov process, then this model is equivalent to the CarHMM with normal emission probabilities. Note that all of the parameter transformations above are one-to-one, so it is easy to go from the CarHMM to the continuous model and back again. This allows for the principled construction of the continuous-time model to be combined with the computational convenience of the CarHMM.

>>>>

Although the CarHMM can incorporate auto-correlation into the structure of an HMM, they break down when observations are taken at irregular time intervals since the probability transition matrix $\Gamma$ would then depend upon the time step between observations. A common solution to this issue is to use a continuous-time method such as the one introduced by Michelot et al \cite{Michelot:2019}, which models the movement of an animal as an Ornstein-Uhlenbeck process with parameters that depend upon the underlying behavioural state of the animal. Namely, a state-switching Ornstein-Uhlenbeck process $Y$ is the solution to the following stochastic differential equation:
%
\begin{equation}
    \label{eqn:SDE}
    dY_t = \beta_{X_t}(\gamma_{X_t} - Y_t)dt + \omega_{X_t} dW_t
\end{equation}
%
where $X_t$ follows some stochastic process which defines the hidden behaviour of the animal at time $t$, $\beta_{X_t}$ relates to rate at which the process returns to its mean value, $\gamma_{X_t}$ is the long-term mean value of the process, $\omega_{X_t}$ is related to short-term variance, and $W$ is a Wiener process. One notable difference between the state-switching OU process and the CarHMM is that the time index $t \in \mathbb{R}$ exists in continuous time and is not necessarily an integer. If the behavioral state $X_t$ is known and does not change between observations, the solution to Equation (\ref{eqn:SDE}) is known to be the following \cite{Michelot:2019}:
%
\begin{align}
Y_{t+\Delta t} | X_{t:t+\Delta t} \sim \mathcal{N}\left((1-e^{-\beta_{X_t}\Delta t})\gamma_{X_t} + e^{-\beta_{X_t}\Delta t} Y_t,\quad \frac{\omega_{X_t}^2}{2\beta_{X_t}} (1-e^{-2\beta_{X_t}\Delta t})\right)
\label{eqn:OU_sol}
\end{align}
%
where $\Delta t$ is the time difference between any two observations $Y_t$ and $Y_{t+\Delta t}$. Continuous time models can be built up from arbitrarily complex stochastic differential equations and allow for uneven step lengths in the observations sequence $Y$. However, most continuous time models require MCMC algorithms to perform inference over and as a result are not easily incorporated into the HHMM structure.

@@

Although most continuous time models are difficult to incorporate into an HHMM, it can be shown that under certain conditions, the CarHMM is equivalent to the state-switching Ornstein-Uhlenbeck process shown in equation (\ref{eqn:OU_sol}). This gives new interpretation to the learned parameters of the CarHMM in the context of a continuous-time model.

\begin{theorem}{Theorem 1.}{}%
If:
\begin{enumerate}
    \item The hidden behavioural process $X$ from (Eq. \ref{eqn:SDE}) follows a Markov chain with $N$ possible states and transitions occurring at equi-spaced time stamps $\left(\Delta t, \ldots, (T-1)\Delta t\right)$, and:
    \item Observations of the SDE from (Eq. \ref{eqn:SDE}) $y_t$ are taken concurrently with every transition of the Markov chain as well as at time $t = 0$, then:
\end{enumerate}
The observations $y$ are equivalent to the output of a conditionally auto-regressive hidden Markov model with normal emission probabilities and parameters $\Theta = (\theta^{(1)}, \ldots, \theta^{(N)}); \enspace \theta^{(s)} = \{\mu^{(s)},\sigma^{2(s)},\phi^{(s)}\}$, where:

\begin{align}
\mu^{(s)} = \gamma_{s}, \qquad \sigma^{2(s)} = \frac{\omega_{s}^2}{2\beta_{s}} (1-e^{-2\beta_{s}\Delta t}), \qquad \phi^{(s)} = e^{-\beta_{s}\Delta t} \label{eqn:CarHMM_to_OU}
\end{align}

\end{theorem}
\begin{proof}{Proof}{}%
Combining equation (\ref{eqn:CarHMM_to_OU}) with equation (\ref{eqn:OU_sol}) gives:
%
\begin{align*}
\left(Y_{i \Delta t} | X_{(i-1)\Delta t: i \Delta t} = s \right) \sim \mathcal{N}\left((1-\phi^{(s)}) \mu^{(s)} + \phi^{(s)} Y_{(i-1) \Delta t}, \enspace \sigma^{2(s)} \right),\\
i = 1, \ldots, T-1
\end{align*}
%
If $X_t$ follows a Markov chain with transitions at the observation times, then the behavioural state $X_t$ does not change between observations and we can re-index $Y_{(i-1) \Delta t} = Y'_{i}$ and $X_{(i-2)\Delta t: (i-1) \Delta t} = X'_i$ for $i = 2,\ldots T$, yielding the desired result:

\begin{align*}
\left(Y'_{i} | X'_{i} = s \right) \sim \mathcal{N}\left((1-\phi^{(s)}) \mu^{(s)} + \phi^{(s)} Y'_{i-1}, \enspace \sigma^{2(s)} \right)\\
i = 2, \ldots, T
\end{align*}
%
$X'$ follows a Markov chain, and the distribution of $(Y'_i|X'_i,Y'_{i+1})$ is consistent with that of a CarHMM with normal emission distributions.
\end{proof}
